---
title: "Section Assignment 6"
date: "2/12/25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(eval = TRUE)
knitr::opts_chunk$set(error = TRUE)
```


***
Things-to-Note: -
1) Don't rename external files or edit them in any way
2) Don't change the variable names
3) Don't destroy or overwrite any variables 
4) Don't use global path in your script. Instead use "setwd()" interactively in the console, but don't include this code in your submission

If you don't take above instructions into account it might effect your grades.
***
Your Information: -
```{r}
firstname <- "Zakir"   # example: "Liam"
lastname <- "Alibhai"  # example: "Mueller"
username <- "zalibhai"  #example: "lomueller"   Do not include the "@ucsd.edu" part!
```



***
Please Read!  

Work carefully this week, as your theory check will also rely on your results from this assignment.


Since we have already downloaded and installed the necessary packages, this week, simply load the tidyverse functions into your working library:
```{r,echo=FALSE}
library(tidyverse)
```
These lines of code above will need to be run each time you start a new R session.

One of the traits of a great programmer is their ability to solve a problem they haven't seen before. One of the best ways to solve a problem you have not seen before is to see if anyone else has. Great programmers are experts of web searching. Do not be afraid to find code on an online help page and run it! 

Copying from the internet is one of the foundations of learning how to program, but it only works as learning if you reflect on why the code you used works. For that reason, this week and in future weeks, you will need to annotate your code or answer a specific question about the process. Use the lines that look like this at the end of each question to input your explanation of the code:

```{r,eval=FALSE,echo=FALSE}
### Your code here
```

 You will be graded on not just your code, but your explanation. Remember, in this class, it is okay to copy code, but you still need to demonstrate independent thought.

  ***  
  

Question 1: Read in this week's data set "Week6SoilData.csv". These data represent inorganic soil nitrogen (NO3-) concentrations from 500 gardens collected in both 2018 and 2021. The nitrogen concentration was determined by combustion analysis and is presented as a proportion of total nitrogen (from 0 to 1). 
```{r}
W6Data<-
head(W6Data)

```


```{r}
### This chunk of code is for grading purposes. 
### Don't change/delete this chunk of code. 
q1 <- W6Data
```

  ***
    


Question 2: How many gardens in 2018 had more than 20% of their total nitrogen in inorganic nitrogen? 
```{r}
High_Inorganic_N<-
High_Inorganic_N

###There are many ways to answer this question. Explain how your code counts how many gardens had high inorganic nitrogen in 2018
# [Your Explanation]
```

```{r}
### This chunk of code is for grading purposes. 
### Don't change/delete this chunk of code. 
q2 <- High_Inorganic_N
```

 ***  
   


Question 3: What type of experimental design would best describe the scenario explained in this assignment? A single word will be sufficient for this answer.

```{r}
q3 <- ""
```

Question 4: Is the Inorganic Nitrogen data normally distributed? Use both a histogram and the ks.test() to support your answer. Think about which column/columns need to be tested.
```{r}
Histogram1<-
Histogram1
```
```{r}
ks.test()
```

```{r}
### This chunk of code is for grading purposes. 
### Don't change/delete this chunk of code. 
q4 <- Histogram1
```

Question 5: Is the Inorganic Nitrogen data we plan to analyze normally distributed? Answer with a TRUE or FALSE statement.

```{r}
q5 <- 
```

  ***  
    


Question 6: Even if you found the data were normally distributed, lets pretend we need to transform it just for practice. Use the `mutate()` function to create a data column that is transformed InorganicN data. Try a square root transformation this time.
```{r}
T_Garden_Data<-
str(T_Garden_Data)

###Explain how you used the mutate function to create a transformed data column
# [Your Explanation Here]
```

```{r}
### This chunk of code is for grading purposes. 
### Don't change/delete this chunk of code. 
q6<- T_Garden_Data
```

  ***  
    

Question 7: Is the transformed Inorganic Nitrogen data normally distributed? Use both a histogram and the ks.test() to support your answer.

```{r}
Histogram2<-
Histogram2
```
```{r}
ks.test()
```

```{r}
### This chunk of code is for grading purposes. 
### Don't change/delete this chunk of code. 
q7<- Histogram2
```

Question 8 (1 point): Is the transformed Inorganic Nitrogen data normally distributed? Answer with a TRUE or FALSE statement

```{r}
q8<- 
```


  ***  

Question 9: Use the appropriate t test to answer the research question: Is there a difference in inorganic nitrogen content between the two years sampled?
```{r}
t.test.result<-

t.test.result
```

```{r}
### This chunk of code is for grading purposes. 
### Don't change/delete this chunk of code. 
q9<- t.test.result
```
    

Question 10: Use ggplot to create a boxplot of the data you used in the t test. Make sure your plot is on the correct column(s) of data. Try to make this plot look professional, like we did in week 3! Hint: `geom_boxplot`
```{r}
Boxplot1<-

Boxplot1

###Explain the steps you took to generate the boxplot. You can do all of that on this line, or add annotation to each line of your code for this question.
# [Your Explanation Here]
```

```{r}
### This chunk of code is for grading purposes. 
### Don't change/delete this chunk of code. 
q9<- Boxplot1
```

  ***  


Once you are done, knit to make sure you have no errors. Then, save the .rmd file with your name and the week number in the file name: 

  (for example: "DONE_Liam_Mueller_Section_Assignment_6.rmd").  
  
Then upload this .rmd file to canvas under the Week 6 module before the deadline. 

You are now more than half way done, look how much programming you have learned in this short time! Congratulations, and keep up the good work. 
