---
title: "Section Assignment 4"
author: "Zakir Alibhai"
date: "1/29/25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(eval = TRUE)
knitr::opts_chunk$set(error = TRUE)
knitr::opts_chunk$set(tidy = TRUE)
```


***
Things-to-Note: -
1) Don't rename external files or edit them in any way
2) Don't change the variable names
3) Don't destroy or overwrite any variables 
4) Don't use global path in your script. Instead use "setwd()" interactively in the console, but don't include this code in your submission

If you don't take above instructions into account it might effect your grades.
***
Your Information: -
```{r}
firstname <- "Zakir"   # example: "Liam"
lastname <- "Alibhai"  # example: "Mueller"
username <- "zalibhai"  #example: "lomueller"   Do not include the "@ucsd.edu" part!
```



***
Please Read!  


This week, your assignment is to work with the 'tidyr' and 'dplyr' packages to rearrange and manipulate data tables. Just like last week the first thing we need to do is turn on the needed packages:
```{r,echo=FALSE}
library(tidyverse)
```

The lines of code above will need to be run each time you start a new R session and need these functions.

Just like last week, one of the traits of a great programmer is their ability to solve a problem they haven't seen before. One of the best ways to solve a problem you have not seen before is to see if anyone else has. Great programmers are experts of web searching. As we are working on becoming expert programmers, this week I have included the link to a set of Rstudio cheat sheets that have high information density all about the different functionalities in the tidyverse. The two most helpful for this week's assignment are the sheets on dplyr and tidyr.

https://rstudio.github.io/cheatsheets/html/tidyr.html

https://rstudio.github.io/cheatsheets/html/data-transformation.html



Copying from the internet is one of the foundations of learning how to program, but it only works as learning if you reflect on why the code you used works. For that reason, this week and in future weeks, you will need to annotate your code. Use the lines that look like this at the end of each question to input your explanation of the code:
```{r,eval=FALSE,echo=TRUE}
### Delete this text and add your own, but don't delete the hashtags (#)
```
 You will be graded on not just your tables, but your explanation. Remember, in this class, it is okay to copy code, but you still need to demonstrate independent thought.

  ***  

Questions:
  
Question 1: The following data table (BadTable) is an untidy display of the population estimate for two different years in three large Texas cities. Use the `pivot_longer()` function to assign to the GoodTable object a table with 3 columns and 6 rows. Hint, `names_to = "Year"` and `values_to = "population"`. For more hints look at the tidyr cheat sheet under "Reshape Data".

```{r}
BadTable<-read.csv(text= "City,2014,2019
                        San Antonio,1436697,1547253
                        Houston,2239558,2320268
                        Austin,912791,978908")

head(BadTable)

GoodTable<-pivot_longer(BadTable, cols = 2:3, names_to = "Year", values_to = "population")

head(GoodTable)

### The pivot_longer function first loads in the variable/vector BadTable, and then selects the years to transform or "tidy" by slicing the second and third columns (in positions 2 and 3 in the csv).  Then, instead of a "wide" format where each row is a city and the years are spread across columns so the populations cells go horizontally, the code makes it "long" so that each row is a city-year combination, where years are in a single column, and cities are repeated so that there can be a separate population column. This makes it so that each row is a different data point, and you can read it vertically. The names and values to create the new 'Year' and population columns so that it's 6 rows for 3 cities by 2 years and 3 columns for City, Year, and Population.
```

```{r}
### This chunk of code is for grading purposes. 
### Don't change/delete this chunk of code. 
q1 <- GoodTable
```

  ***  
  Question 2: Why is GoodTable a better format than BadTable? Put another way, what makes a table tidy?
  
```{r}
q2 <-  "The GoodTable 'long' format is better than the 'wide' BadTable because each variable now has its own column to compare, rather than the intersecting CSV of 'CityxYear=population number' with multiple populations per row. Now each instance of data, or the measurements of populations, have their own rows, and aren't spread across the columns. This is labelled better and it's easier to see at a glance the corresponding numbers by city and year. It is 'tidy' because it is easier to read with a separate label for years and the population, which is probably easier for axis graphing too.
The documentation says that each variable is its own column, each observation, or case of population, is in its own row, and you can access variables as vectors of columns due to the way R works."
```

  ***  
  
Question 3: Extract the rows from "GoodTable" where population is less than 1500000 and save them to a the object "SmallerCities" Hint: Page 1 of the dplyr cheat sheet, "Manipulate Cases"
```{r}
SmallerCities<-GoodTable |> filter(population < 1500000)
head(SmallerCities)
### This does exactly what the documentation says: it loads GoodTable, uses the dplyr 'pipes' syntax (like squeezing the data) to call a filter function, and applies it and extracts the rows with our desired values by evaluating the conditional of cell / case values where the population is less than 1500000. If true, it extracts that column; kind of like an automatic for loop.
```

```{r}
### This chunk of code is for grading purposes. 
### Don't change/delete this chunk of code. 
q3 <- SmallerCities
```


  ***  
  
Question 4: Rearrange "GoodTable" by high to low population. Hint: Page 1 of the dplyr cheat sheet, "Manipulate Cases"
```{r}
OrderedGoodTable<-GoodTable |> arrange(desc(population))
head(OrderedGoodTable)
### Once again, this does exactly what the documentation says, and calls the 'arrange' function on GoodTable to manipulate it by literally just rearranging it according to a variable I chose, in this case population, and we add the 'desc' function call to order it high to low, or in 'descending' order, so that the table is ordered this way.
```

```{r}
### This chunk of code is for grading purposes. 
### Don't change/delete this chunk of code. 
q4 <- OrderedGoodTable
```


  ***  

Question 5 & 6: To really get a feel for the tidyverse, we need to start working with bigger data sets. The following two data tables are subsets from the flights and weather data in the "nycflights2013" package. Download The "flights.csv" and "JFKWeather.csv" files from this week's section activity on Canvas. "flights.csv" contains flight information for the 28 Hawaiian Airlines flights that departed JFK airport in February 2013. The "JFKWeather.csv" contains hourly weather data from JFK airports for the month of February 2013. Read these .csv files into R below:


```{r}
flights<-read.csv(file = "flights.csv")

weather<-read.csv(file = "JFKWeather.csv")

```

```{r}
### This chunk of code is for grading purposes. 
### Don't change/delete this chunk of code. 
q5 <- flights
q6 <- weather
```


  ***  

Question 7: Our goal is to combine the flight and weather data tables into one large table containing the flight information and the weather data captured the hour each flight was scheduled to take off. If you do this correctly, the resulting table should have exactly 28 rows and 23 columns. Hint: `inner_join()` on page 2 of the dplyr cheat sheet and the sections titled "RELATIONAL DATA" and "COLUMN MATCHING FOR JOINS". Hint 2: Both of these data tables share the "time_hour" column. 

```{r}
CombinedTable<-inner_join(flights, weather, by = join_by(time_hour))
str(CombinedTable)
### Here, I called the inner_join with the syntax in the documentation, specifying the flights and weather datasets as my x and y variables to be combined or 'joined'. This function explicitly joins or concatenates data, which is what I want with adding 11 more columns (since each have 12, with the last one, time_hour, being a repeat/match or overlapping column) of the correct number of rows, as it only keeps the rows with matches. This is good because the instructions say both data tables share the 'time_hour' column, and thus matched the values since they're in both. I used column matching with 'join_by' so that it looked at the 'time_hour' column, and then for each flight, it found the weather data from the same hour, and then it merged the matching rows together, and got rid of any rows that didn't match. So basically, it matched each flight with the weather data at its time, again like a more complex for loop.
```

```{r}
### This chunk of code is for grading purposes. 
### Don't change/delete this chunk of code. 
q7 <- CombinedTable
```

   ***   
   
Question 8: What type of data is the third column of your dataset? Answer with only a single word that descibes the type of data. ("Continuous", "Count", "Ordinal", "Binomial", "Categorical")

```{r}
q8<-"Continuous"
```


 ***  
 
Question 9: Let's say, instead of the weather at the time of the flight, we want to know the average temperature each day in February 2013, at JFK. You will need to combine the `summarise()` and `group_by()` functions with the "weather" data you created back in question 4 to achieve a data table with 28 rows and two columns (day and average temp). Hint: Look at the left column on the first page of the dplyr cheat sheet. Hint 2: Your answer will use "|>" twice. These are called pipes and allow you to perform sequential operations within a single command. Essentially, "|>" acts like "and then do". For example: "A |> B |> C" really means "take A and then do B to it and then do C to it". 

```{r}
MeanDailyTemp<- weather |> group_by(day) |> summarise(avg = mean(temp))
  
str(MeanDailyTemp)
### Here, we use the 'pipes' to perform sequential operations on the weather dataset; first, I used the 'group_by' as shown in the documentation, before 'summarise', to create a copy of the table that groups all of the data by the 'day' column, so that when I call 'summarize', it will 'manipulate' each 'group' or day separately, and then combine the results. For the summarize part, I again followed the syntax in the documentation, which takes the column of temperatures 'temp' as the input vector, and then finds the mean of each case or row and returns the average back. It then stores the average of all the temperature values within each day group (since it was grouped/sorted that way) into a new column, 'avg', which is the average temperature for each day. Once again, this is a lot like a one-line for-loop that performs the mean operation on temperatures for each day and stores it into the new dataframe with just columns for the day and avg.
```

```{r}
### This chunk of code is for grading purposes. 
### Don't change/delete this chunk of code. 
q9 <- MeanDailyTemp
```

 ***  
Question 10 & 11: Calculate the mean and median for the DailyTemp data you just generated.
```{r}
meanfebtemp <-mean(MeanDailyTemp$avg)
meanfebtemp

medianfebtemp <-median(MeanDailyTemp$avg)
medianfebtemp

### In which direction are the data skewed?
```

```{r}
### This chunk of code is for grading purposes. 
### Don't change/delete this chunk of code. 
q10<- meanfebtemp
q11<- medianfebtemp
```

***

Question 12: In which direction are the data skewed? Answer with only a single phrase: ("Left", "Right")

```{r}
q12 <- "Right"
```


Once you are done, click the "Knit" button above(It looks like a blue ball of yarn) to check your work. Then download the rmd file with your name and the week number in the file name: 

  (for example: "Liam_Mueller_Section_Assignment_4.rmd").  
  
Then upload this rmd file to canvas under the Week 4 section assignment before the deadline. 

And that is it for this week! Keep up the amazing work!
